---
title: "HappinessAnalysis"
author: "TomasVarga"
slug: happiness
date: "10/20/2020"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(maps)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(scales)
library(countrycode)

data <- read.csv("2017HappinessAnalysis.csv", sep=",")
```

# Happiness around the world

I always enjoyed geography and comparing countries on various metrics. The most interesting metric might be happiness which itself depends on many other variables. I used the world happiness dataset from Kaggle to analyse the connection between happiness and various factors around the world.

For more information check out the original dataset: https://www.kaggle.com/unsdsn/world-happiness

## Family and happiness

```{r}
#We create histograms for each continent's happiness based on family score in countries
data %>% 
  mutate(
    continent = countrycode(Country, 'country.name', 'continent')
  ) %>% 
  group_by(continent) %>% 
  summarise(
    Family = Family,
    mean_fam = mean(Family)
  ) %>% 
  na.omit() %>% 
  ggplot(aes(x=Family)) +
  geom_histogram() +
  facet_wrap(~continent, scales="fixed", ncol=1) +
  theme_bw() +
  labs(title="Family plays a crucial role in people's happiness",
       subtitle = "Family score across regions") +
  ylab("") +
  xlab("How important is family for your happiness?")
```

## Wealth and happiness

```{r}
#We create scatterplot with trendline to study relationship
data %>% 
  ggplot(aes(x=Economy..GDP.per.Capita., y=Happiness.Score)) +
  geom_point() +
  geom_smooth(method="lm",se=FALSE)+
  theme_bw() +
  labs(title = "Wealthier contries are happier",
       subtitle="Visible trend between happiness and GDP per capita") +
  ylab("Happiness score") +
  xlab("GDP per capita proxy")
```

##How does happiness look like on a map?

```{r}
#We create world map with countries coloured based on happiness score

world <- ne_countries(scale = "medium", returnclass = "sf")

#rename countries so that they match
world$sovereignt <- gsub("United States of America", "United States", world$sovereignt)

#we merge data and want to keep all countries on map (hence last part of line)
world <- merge(world, data, by.x = "sovereignt", by.y = "Country", all.x = T)

ggplot(data = world) +
  geom_sf(aes(fill = Happiness.Score), lwd = 0) +
  scale_fill_gradient(low=rgb(253, 241, 240, maxColorValue = 255), high=rgb(224, 56, 56, maxColorValue = 255),
                      limits = c(3, 7), na.value = rgb(223, 230, 233, maxColorValue = 255), oob=squish) +
  theme_minimal() +
  labs(fill = "Happy") +
  theme(plot.margin = unit(c(.05,.05,.05,.05), "lines")) +
  labs(title ="More developed countries scored higher",
       subtitle="Happiness score around the world")
```

## Modelling Happiness
```{r}

Happiness_model<-lm(Happiness.Score~Family +
     Health..Life.Expectancy. + 
     Economy..GDP.per.Capita.+ 
     Freedom+
     Generosity+
     Trust..Government.Corruption.
    ,
       
   data=data)

Happiness_model%>%broom::tidy()
Happiness_model%>%broom::glance() 

Happiness_model

```
We can see that proxies for health, wealth, freedom, generosity and trust in government all positively influence the happiness of a country and altogether explain 80.5% of the variation in happiness.